### 2018-04-11 paper.Rmd
intro

abstract: |
  The Child Language Data Exchange System (CHILDES) has played a critical role in research on child language development. However, current methods of using CHILDES data rely on a command line interface to extract developmental indices or custom parsers to extract data for larger-scale analyses. We introduce childes-db, a database-formatted mirror of CHILDES that improves data accessibility and usability by offering novel interfaces, including browsable web applications and an R application programming interface (API). Along with a versioned infrastructure that facilitates reproducibility of past analyses, these interfaces lower barriers to analyzing naturalistic parent-child language, allowing for a wider range of researchers in language and cognitive development to easily leverage CHILDES in their work.

@@

Children's language acquisition is a core puzzle in cognitive science. How are the representations underlying the generative use of language learned over children's first years? Because a language is comprised of many individual pieces of information -- the names of particular objects, the marker for the past tense, the position of determiners -- language development is fundamentally a question of learning from environmental input, which is the only source of information on these specifics. Further, children are often reluctant experimental participants; often their most revealing and most sophisticated uses of language emerge in naturalistic play. Thus, language acquisition research must be grounded in the ecological behavior of children, both their language input and their language production in context.

The foundational resource for studies of language acquisition is recordings and transcripts of children. Starting with Roger Brown's [-@brown1973] work on Adam, Eve, and Sarah, longitudinal recordings of children's talk (and the talk they hear from their caregivers) have formed the basis for this field. But creating individual transcripts is extremely hard work. Beyond gathering the data, even orthographic transcriptions (without phonetic detail) can take ten times the length of the original recording to transcribe. Thus, researchers' desire for data from multiple ages or multiple children can very quickly outstrip their resources.

@@

Yet there are some challenges working with CHILDES, both for students and for advanced users. The CHILDES ecosystem uses a special file format (CHAT), which is stored in a plain text format but includes structured annotations grouped into tiers (stored on separate lines) so that information about utterances can be stored with accompanying information such as the phonological or morphological structure of the utterance. These files are analyzed using a command-line tool (CLAN) with a host of specific arguments that allow users to count word frequencies, compute statistics (e.g., mean length of utterance, or MLU). While this system is flexible and powerful, mastering the CHAT codes and the CLAN tool can be daunting to navigate for a beginner or student. These technical barriers decrease the ease of casual exploration by a novice researcher or in a classroom exercise.

On the opposite end of the spectrum, for data-oriented researchers who are interested in doing large-scale analyses of CHILDES, the current tools are also not ideal. CLAN software is an excellent tool for interactive exploration, but -- as a free-standing application -- can be tricky to build into a processing pipeline written in a modern analytic platform such as Python or R. Thus, researchers who would like to ingest the entire corpus (or some large subset) into a computational analysis typically write their own parsers of the CHAT format to extract the subset of the data they would like to use [e.g., @redington1998;@yang2013;@meylan2017].

This practice is problematic for a number of reasons. Effort is wasted in rewriting the same parser again and again. In addition, this process introduces a major source of errors and inconsistencies in data handling due to difficulties dealing with the many special cases in the CHAT standard. Further, these scripts are rarely shared, leading to much greater difficulty in reproducing the exact numerical results from papers using CHILDES [see e.g., @meylan2017]. This lack of computational reproducibility is widely recognized as problematic because it does not allow researchers to verify or build on published research [@donoho2010;@stodden2016]. Finally, the CHILDES corpus itself is a moving target;  computational work using the entire corpus at one time point may include a different set of data than subsequent work (due to new contributions or revisions).

>>>>

abstract: |
  The Child Language Data Exchange System (CHILDES) has played a critical role in research on child language development. Access to these data can be complex for novices and difficult to automate for advanced users, however. To address these issues, we introduce childes-db, a database-formatted mirror of CHILDES that improves data accessibility and usability by offering novel interfaces, including browsable web applications and an R application programming interface (API). Along with a versioned infrastructure that facilitates reproducibility of past analyses, these interfaces lower barriers to analyzing naturalistic parent-child language, allowing for a wider range of researchers in language and cognitive development to easily leverage CHILDES in their work.

@@

Children's language acquisition is a core puzzle in cognitive science. How are the representations underlying the productive use of language learned over children's first years? Because a language is comprised of many individual pieces of information -- the names of particular objects, the marker for the past tense, the position of determiners -- language development is fundamentally a question of learning from environmental input, which is the only source of information on these specifics. Further, children are often reluctant experimental participants; often their most revealing and most sophisticated uses of language emerge in naturalistic play. Thus, language acquisition research must be grounded in the ecological behavior of children, both their language input and their language production in context.

The foundational resource for studies of language acquisition is recordings and transcripts of children. Starting with Roger Brown's [-@brown1973] work on Adam, Eve, and Sarah, longitudinal recordings of children's talk (and the talk they hear from their caregivers) have formed the basis for this field. But creating individual transcripts is extremely hard work. Beyond gathering the data, even orthographic transcriptions (without phonetic detail) can take ten times the length of the original recording to transcribe [@macwhinney2000]. Thus, researchers' desire for data from multiple ages or multiple children can very quickly outstrip their resources.

@@

Yet there are some challenges working with CHILDES, both for students and for advanced users. The CHILDES ecosystem uses a special file format (CHAT), which is stored as plain text but includes structured annotations grouped into tiers (stored on separate lines) so that information about utterances can be stored with accompanying information such as the phonological or morphological structure of the utterance. These files are analyzed using a command-line tool (CLAN) with a host of specific arguments that allow users to count word frequencies, compute statistics (e.g., mean length of utterance, or MLU). While this system is flexible and powerful, mastering the CHAT codes and the CLAN tool can be daunting for a beginner or student. These technical barriers decrease the ease of casual exploration by a novice researcher or in a classroom exercise.

On the opposite end of the spectrum, for data-oriented researchers who are interested in doing large-scale analyses of CHILDES, the current tools are also not ideal. CLAN software is an excellent tool for interactive exploration, but -- as a free-standing application -- it can be tricky to build into a processing pipeline written in a modern analytic language such as Python or R. Thus, researchers who would like to ingest the entire corpus (or some large subset) into a computational analyses typically write their own parsers of the CHAT format to extract the subset of the data they would like to use [e.g., @redington1998;@yang2013;@meylan2017].

The practice of writing custom parsers is problematic for a number of reasons. Effort is wasted in rewriting the same parser again and again. In addition, this process introduces a major source of errors and inconsistencies in data handling due to difficulties dealing with the many special cases in the CHAT standard. Further, these parsing scripts are rarely shared, leading to much greater difficulty in reproducing the exact numerical results from papers using CHILDES [see e.g., @meylan2017 for an example]. In addition, the CHILDES corpus itself is a moving target; computational work using the entire corpus at one time point may include a different set of data than subsequent work (due to new contributions or revisions) and there is no simple way for researchers to document exactly what version of the corpus has been used short of creating a full mirror of the data. These factors together lead to a lack of computational reproducibility, a major problem that keeps researchers from verifying or building on published research [@donoho2010;@stodden2016].
